Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-05-30 schemes.tex
Attempt to merge

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding,
all the original data packets $P_j, j \in [1,g]$, each of $B$ bits, are used
to create coded packets. In the following subsections, we describe the
basic functionalities of \ac{RLNC}. Later, we mention applications that could
potentially benefit from including \ac{RLNC} as a coding scheme.

\subsubsection{Encoding}
With \ac{RLNC}, each coded packet is a random linear
combination of the original set of packets. Hence,
a \ac{l.i.} set of $g$ coded packets, $C_i,\ i \in [1,g]$
is required in order to get the original information.
Each original packet is considered as a concatenation of elements
from a \ac{GF} of a given size $q$, which we denote
$GF(q)$. To create a coded packet, a coding coefficient
$v_{i,j}$, is chosen at random from $GF(q)$ for every
packet $P_j$ and multiplied and added
following the respective \ac{GF} arithmetics. In this way,
a coded packet is:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{i,j} \otimes P_j ,\ \forall i \in [1,g]
\end{align}

To indicate which packets were used to generate a coded packet, one
form is to append its coding coefficients. In this case, the overhead
included for $C_i,\ \forall i \in [1,g]$ by the coding coefficients
is given by:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{i,j}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
To perform decoding, we define $\textbf{C} = \left[C_1 \ldots C_g \right]^T$
and $\textbf{P} = \left[P_1 \ldots P_g \right]^T$. Then, decoding
reduces to solve the linear system $\textbf{C} = \textbf{V} \cdot \textbf{P}$
using Gaussian elimination \cite{fragouli2006network}.
Here, the coding matrix $\textbf{V}$ contains any set of $g$ linearly
independent packets $C_i$ as rows as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{1,1} & \ldots & v_{1,g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g,1} & \ldots & v_{g,g} \\
\end{array}
\right]
\end{align}
%
The decoder begins to compute and remove the contributions from each
of the pivot elements, e.g. leftmost elements in the main diagonal of
\eqref{eq:coding_matrix}, to reduce $\textbf{V}$ to reduced echelon
form. In this way, it is possible to recover the original set of packets.

>>>>

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding,
all the original data packets $P_j, j \in [1,g]$, each of $B$ bits, are used
to create coded packets. In the following subsections, we describe the
basic functionalities of \ac{RLNC}.

\subsubsection{Encoding}
In \ac{RLNC}, any coded packet is a linear combination from all
the original packets. For the coding scheme, packets are seen as
algebraic entities formed as a sequence of elements from $GF(q)$,
which is a \ac{GF} of size $q$. Later, each original packet is
multiplied by a coding coefficient from $GF(q)$. The coding coefficients
are picked uniformly at random by the encoder. To perform the
multiplication of a packet by a coding coefficient, the coefficient is
multiplied for each of the elements in the concatenation that composes
an original packet, preserving the concatenation. Later, all resulting
packets are added within the \ac{GF} arithmetics together to generate
a coded packet. Thus, a coded packet can be written as:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{i,j} \otimes P_j ,\ \forall i \in [1,\ldots)
\end{align}

In \eqref{eq:coded_packet}, $C_i$ is the generic coded packet. In principle,
the encoder may

For signaling a receiver which of the original packets were utilized to
create a coded packet, a simple choice is to append its coding coefficients
as a header of the coded packet. Hence, an amount of overhead is included
since every coded packet given that we need to provide some signalling.
The coding coefficients overhead can be quantied as:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{i,j}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
To be able to decode a batch of $g$ packets, a \ac{l.i.} set of $g$
coded packets, $C_i,\ i \in [1,g]$ is required in order to get the
original information. Once this set has been collected for any receiver,
the original packets can be found by computing the solution of a system
of linear equation using \ac{GF} aritmethics. Thus, we define
$\textbf{C} = \left[C_1 \ldots C_g \right]^T$ and
$\textbf{P} = \left[P_1 \ldots P_g \right]^T$. Algebraically, decoding
simplifies to finding the inverse of $\textbf{V}$ in the linear system
$\textbf{C} = \textbf{V} \cdot \textbf{P}$, which can achieved using
Gaussian elimination techniques \cite{fragouli2006network} that can
be efficient. For decoding to be possible, the (coding) matrix $\textbf{V}$
must contain any set of $g$ linearly independent packets $C_i$ in order
to reduce to row-echelon form. The coding matrix is represented as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{1,1} & \ldots & v_{1,g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g,1} & \ldots & v_{g,g} \\
\end{array}
\right]
\end{align}
%
The decoder begins to compute and remove the contributions from each
of the pivot elements, e.g. leftmost elements in the main diagonal of
\eqref{eq:coding_matrix}, to reduce $\textbf{V}$ to reduced echelon
form. In this way, it is possible to recover the original set of packets.
